% =====================================================================
%  Optimize Against EA — API Documentation (Developer Reference)
%  GameLab 2 (2025) — Julius-Maximilians-Universität Würzburg
%  Authors: Philipp Sehne, David Ahrens
%
%  Build:  pdflatex api.tex   (run twice for TOC/links)
%
%  LAYOUT NOTE: This file is the structural skeleton of the API doc.
%  A few entries are fully written out (marked "worked example") to fix
%  the format; the rest are \todo stubs following the same pattern.
%  Every entry uses the \apientry environment defined below.
% =====================================================================
\documentclass[11pt,a4paper]{report}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[margin=2.5cm]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[colorlinks=true, linkcolor=blue!50!black, urlcolor=blue!50!black]{hyperref}

% --- TypeScript listings style ---------------------------------------
\definecolor{codebg}{gray}{0.96}
\definecolor{codekw}{RGB}{0,0,180}
\definecolor{codestr}{RGB}{160,50,20}
\definecolor{codecomment}{RGB}{60,120,60}
\lstdefinelanguage{TypeScript}{
  keywords={import, export, from, type, interface, const, let, function,
            return, if, else, for, while, new, extends, implements,
            readonly, number, string, boolean, void, null, undefined,
            true, false, as, satisfies},
  keywordstyle=\color{codekw}\bfseries,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{codecomment}\itshape,
  string=[b]',
  morestring=[b]",
  morestring=[b]`,
  stringstyle=\color{codestr},
  sensitive=true,
}
\lstset{
  language=TypeScript,
  basicstyle=\small\ttfamily,
  backgroundcolor=\color{codebg},
  frame=single, framerule=0pt,
  breaklines=true, breakatwhitespace=true,
  columns=fullflexible, keepspaces=true,
  aboveskip=1ex, belowskip=1ex,
  literate={↑}{{$\uparrow$}}1 {→}{{$\rightarrow$}}1
           {↓}{{$\downarrow$}}1 {←}{{$\leftarrow$}}1,
}

% --- helpers ----------------------------------------------------------
\newcommand{\todo}[1]{\textcolor{red!70!black}{\textbf{[TODO: #1]}}}
\newcommand{\srcfile}[1]{\texttt{\small #1}}   % source path
\newcommand{\ts}[1]{\texttt{#1}}               % inline identifier

% One API reference entry: \begin{apientry}{name}{source file} ... \end{apientry}
\newenvironment{apientry}[2]{%
  \par\medskip\noindent
  \begin{minipage}{\textwidth}
  \hrule\vspace{0.6ex}
  {\large\ttfamily\bfseries #1}\hfill{\footnotesize\srcfile{#2}}
  \vspace{0.4ex}\hrule\medskip
}{%
  \end{minipage}\par\medskip
}

% Labelled paragraph inside an entry: Params / Returns / Side effects / Notes
\newcommand{\apifield}[1]{\par\smallskip\noindent\textbf{#1.}\ }

\title{\Huge\textbf{Optimize Against EA}\\[1ex]
       \LARGE API Documentation\\[0.5ex]
       \large Developer Reference\\[2ex]
       \large GameLab~2 --- Julius-Maximilians-Universität Würzburg}
\author{Philipp Sehne \\ \small\texttt{Philipp.Sehne@stud-mail.uni-wuerzburg.de}
   \and David Ahrens  \\ \small\texttt{David.Ahrens@stud-mail.uni-wuerzburg.de}}
\date{\today}

\begin{document}

\maketitle
\tableofcontents

% =====================================================================
\chapter{Introduction}
% =====================================================================

\section{Scope}

\emph{Optimize Against EA} is a purely client-side React~19 + TypeScript
single-page application (built with Vite). It exposes no HTTP/REST endpoints;
``API'' in this document therefore means the \textbf{programmatic interfaces
of the codebase}: the exported types, engine functions, Web Worker message
protocols, React hooks, and data serialization formats a developer needs in
order to extend the application or reuse its parts.

This document complements the \emph{Handbook \& User Manual}, which covers
the application from the player's perspective.

\section{How to read a reference entry}

Each documented symbol appears as an entry with its source file in the
header, followed by its signature, a description, and where relevant
\textbf{Params}, \textbf{Returns}, \textbf{Side effects} and \textbf{Notes}
fields. All paths are relative to
\srcfile{OptimizeAgainstEA/OptimizeAgainstEAApp/src/}.

\section{Conventions used in the codebase}

\begin{itemize}
  \item Type-only imports use \ts{import type \{...\}}.
  \item React components are \ts{function} declarations; engine and utility
        functions are \ts{const} arrow functions or plain functions.
  \item No \ts{any}; all callback parameters are explicitly typed.
  \item Fitness in the map/maze EAs is \emph{minimized} (0 = optimal);
        fitness in the shooter GA is \emph{maximized}.
  \item Randomness is injected as \ts{RNG = () => number} (a
        \ts{Math.random}-compatible function) so engines stay deterministic
        under a seeded RNG.
\end{itemize}

% =====================================================================
\chapter{Architecture Overview}
% =====================================================================

\section{Layering}

Every game module follows the same four-layer structure. Lower layers never
import from higher ones:

\begin{center}
\begin{tabularx}{\textwidth}{l X}
  \toprule
  \textbf{Layer} & \textbf{Contents} \\
  \midrule
  \ts{types/}      & Pure type contracts and default configurations. No logic. \\
  \ts{engine/}     & Pure, React-free logic: problem construction, EA
                     operators, codecs, geometry. Unit-testable in isolation. \\
  \ts{engine/*.worker.ts} & Web Workers running the EA off the main thread;
                     communicate via typed message unions
                     (\ts{WorkerInMessage}/\ts{WorkerOutMessage}). \\
  \ts{hooks/}      & React hooks bridging engine/workers to component state. \\
  \ts{components/} & Presentational and interactive React components. \\
  \bottomrule
\end{tabularx}
\end{center}

\section{Module map}

\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  \textbf{Module} & \textbf{Route} & \textbf{Chapter} \\
  \midrule
  \srcfile{modules/BattleShips/}  & \ts{/PeakFinder}   & \ref{ch:api-battleships} \\
  \srcfile{modules/mazeGame/}     & \ts{/MazeExplorer} & \ref{ch:api-maze} \\
  \srcfile{modules/shooterGame/}  & \ts{/ShooterGame}, \ts{/HordeGame} & \ref{ch:api-shooter} \\
  \srcfile{components/, hooks/, context/} & --- (shared) & \ref{ch:api-shared} \\
  \bottomrule
\end{tabularx}
\end{center}

\todo{Architecture diagram: modules, workers, stores, and the data flow
between them (one figure).}

% =====================================================================
\chapter{Shared Infrastructure}
% =====================================================================
\label{ch:api-shared}

\section{Global state and settings}

\begin{apientry}{SettingsContext / SettingsProvider}{context/SettingsContext.tsx}
Global settings provider; wraps all routes in \srcfile{App.tsx}. Holds
three independent settings objects, each with its own setter:
\begin{lstlisting}
const { eaSettings,      setEaSettings,        // EASettings
        shooterSettings, setShooterSettings,   // ShooterSettings
        hordeSettings,   setHordeSettings,     // HordeSettings
      } = useSettings(); // throws outside a <SettingsProvider>
\end{lstlisting}
\ts{EASettings} tunes the Solo-Play shooter GA (mutation rate/strength,
presim generations, population size, crossover type, hall of fame,
analytics retention, injection deviation). \ts{ShooterSettings} holds
starter DNA, round duration, tug-of-war win threshold, player stats and
the powerup-choice options. \ts{HordeSettings} is deliberately separate
from \ts{EASettings} (own difficulty presets, own starter DNA padded with
neutral 0.5s for the Horde-only genes, wave sizing, map id, custom map
layout, kills-per-upgrade).
\apifield{Notes}State is React-only --- \emph{not} persisted; a reload
restores the \ts{default*Settings}. \ts{resetEASettings()} /
\ts{resetShooterSettings()} / \ts{resetHordeSettings()} return fresh
defaults (the shooter/horde variants reset the starter DNA to all-0.1
``untrained'' genes).
\end{apientry}

\section{Layout components}

\begin{apientry}{GameLayout, LAYOUT}{components/layout/GameLayout.tsx}
Three-column page shell (\ts{leftBar | canvas | sidebar}) used by the
shooter pages. Exports the \ts{LAYOUT} constant (column widths, spacing,
panel colours) for canvas-sizing arithmetic.
\apifield{Notes}Pass the combined sidebar width to \ts{useScaledCanvas}
so the canvas scales correctly.
\end{apientry}

\begin{apientry}{PageContainer}{components/layout/PageContainer.tsx}
\ts{PageContainer(\{ children, backgroundImage? \})} --- full-viewport
page wrapper (\ts{.page-container}); \ts{backgroundImage} is a URL or
imported image rendered as a covering, centered background. Default
export.
\end{apientry}

\section{Shared hooks}

\begin{apientry}{useScaledCanvas (worked example)}{hooks/useScaledCanvas.ts}
\begin{lstlisting}
const { containerRef, scale } = useScaledCanvas({
  baseWidth:    number,   // logical canvas width  (e.g. ARENA.WIDTH)
  baseHeight:   number,   // logical canvas height
  sidebarWidth: number,   // total horizontal space taken by side panels
  padding:      number,   // default 16
});
\end{lstlisting}
Computes the largest uniform scale at which the logical canvas fits the
measured container minus \ts{sidebarWidth} and \ts{padding}, preserving
the aspect ratio.
\apifield{Returns}\ts{containerRef} (attach to the element whose size to
measure --- a \ts{ResizeObserver} keeps \ts{scale} current) and the
numeric \ts{scale} factor to apply to the canvas.
\end{apientry}

\begin{apientry}{useViewport}{hooks/useViewport.ts}
\ts{useViewport(): \{ W, H \}} --- the current
\ts{window.innerWidth}/\ts{innerHeight}, re-rendered on every window
\ts{resize}.
\end{apientry}

\begin{apientry}{useOrientationLock}{hooks/useOrientationLock.ts}
\ts{useOrientationLock(type = 'landscape')} --- locks the screen
orientation while mounted, unlocking on unmount. Only attempts the lock
when fullscreen is active (a platform requirement) and re-locks on every
\ts{fullscreenchange}. Fails silently where unsupported (iOS Safari);
works on Android Chrome in fullscreen.
\end{apientry}

\begin{apientry}{useReplayPlayer}{hooks/useReplayPlayer.ts}
\begin{lstlisting}
const { frameIndex, currentFrame, totalFrames, isPlaying,
        isFirst, isLast, next, prev, goTo, play, pause } =
  useReplayPlayer<Frame>(frames, autoplayIntervalMs = 1200);
\end{lstlisting}
Playback state machine for any recorded frame list --- shared by the
PeakFinder EA replay, the Maze EA replay and the generation replay.
Generic over the frame type: it only navigates, never looks inside a
frame. \ts{play(ms?)} auto-advances at the given dwell (default
\ts{autoplayIntervalMs}) and self-terminates on the last frame;
\ts{next}/\ts{prev}/\ts{goTo} clamp to the valid range.
\end{apientry}

\begin{apientry}{useSavedLibrary}{hooks/useSavedLibrary.ts}
\begin{lstlisting}
const { entries, save, remove, rename } =
  useSavedLibrary<Input, Entry extends SavedEntryBase>(storageKey, toEntry);
interface SavedEntryBase { id: string; name?: string; createdAt: number }
\end{lstlisting}
localStorage-persisted library of player creations --- the shared core
behind \ts{useSavedMaps} (\ts{bs.savedMaps}) and \ts{useSavedMazes}
(\ts{maze.savedMazes}). \ts{toEntry} converts the game's input into an
entry; its \ts{id} de-duplicates (saving an existing id is a no-op, and
\ts{save} returns the id either way). A blank name in
\ts{save}/\ts{rename} clears the name. Storage access is guarded, so
private-mode failures degrade to an in-memory list.
\end{apientry}

\section{Hint system}

Page-wide, toggleable contextual help
(\srcfile{components/hints/}). Content lives exclusively in
\srcfile{hintContent.ts} (BattleShips) and
\srcfile{modules/mazeGame/hints/mazeHintContent.ts}.

\begin{apientry}{useHints (worked example)}{components/hints/HintContext.tsx}
\begin{lstlisting}
const { showHint, dismiss } = useHints();
showHint('vsEa.afterProbe', { vars: { gens: '5' }, actions: [...] });
\end{lstlisting}
Fires a hint by id. No-ops when hints are disabled or the hint is marked
\ts{once} and was already seen this session.
\apifield{Side effects}Persists \ts{enabled} to \ts{localStorage}
(\ts{bs.hints.enabled}) and \ts{seen} ids to \ts{sessionStorage}
(\ts{bs.hints.seen}).
\end{apientry}

\begin{apientry}{HintDef}{components/hints/hintContent.ts}
\begin{lstlisting}
{ title?, body, style: 'modal' | 'toast', once?, pauses? }
\end{lstlisting}
\ts{pauses: true} blocks interaction behind a backdrop until dismissed.
\end{apientry}

\begin{apientry}{HintPopover}{components/hints/HintPopover.tsx}
Anchored, non-blocking coach-mark: wrap any element to pin a hint bubble
to it.
\begin{lstlisting}
<HintPopover id="vsEa.replayButton" placement="top" show={canReplay}>
  <button>Watch Last Replay</button>
</HintPopover>
\end{lstlisting}
Props: \ts{id} (a \ts{HintId} from \ts{hintContent.ts} --- the wording
lives there), \ts{children} (the anchor), \ts{placement}
(\ts{'top' | 'top-start' | 'top-end' | 'bottom' | 'left' | 'right'},
default \ts{'right'}), \ts{show} (extra condition on top of
enabled/once-seen, default \ts{true}), \ts{vars} (fills
\ts{\{placeholders\}} in the body), \ts{actions} (buttons; falls back to
a single ``Got it''), \ts{dismissAfter} (auto-dismiss in ms). The bubble
renders via a portal to \ts{document.body} and is clamped to stay fully
on-screen.
\apifield{Notes}CSS-grid caveat: the wrapper is \ts{display:
inline-block}, so inside a grid it would itself become the grid item ---
wrap the child \emph{inside} the grid item instead.
\end{apientry}

\section{Help system}

\begin{apientry}{HELP\_TOPICS / HelpButton}{components/help/helpContent.tsx}
Content registry for the site-wide help modals. Adding a topic = one entry
in \ts{HELP\_TOPICS} plus a \ts{<HelpButton topic="..." />}. Each topic
provides a \ts{Gameplay} and a \ts{Technical} tab component.
\end{apientry}

\section{Firebase}

\begin{apientry}{firebase}{lib/firebase.ts}
Exports \ts{firebaseApp} (the initialized Firebase app) and \ts{db} (its
Firestore handle). The client-side web config is hardcoded in this file
(project \ts{optimizeagainstea}); as a public web API key it is not a
secret --- access control lives in the Firestore security rules.
\apifield{Notes}The only consumer is the community Raidboss mode
(\srcfile{modules/shooterGame/game/raidbossStore.ts}); every other
feature is fully client-local. Swapping the backing project means
editing this one config object.
\end{apientry}

\section{Utilities}

\begin{apientry}{rng}{utils/rng.ts}
\begin{lstlisting}
type RNG = () => number;          // next value in [0, 1)
makeLCG(seed?: number): RNG
\end{lstlisting}
Seeded linear congruential generator (\ts{s = 1664525 s + 1013904223}
mod $2^{32}$, via \ts{Math.imul} --- pure integer math, so the stream is
identical on every device). The single random source of the PeakFinder
and Maze EAs and of maze generation: same seed $\Rightarrow$ fully
reproducible run. Without a seed, a fresh \ts{Math.random()}-derived one
is used; the degenerate all-zero state is remapped.
\end{apientry}

\begin{apientry}{listenable}{utils/listenable.ts}
\begin{lstlisting}
createListenable(): { notify(): void;
                      subscribe(fn: () => void): () => void }
\end{lstlisting}
The observer plumbing behind every module-level singleton store
(\ts{gameStore}, \ts{analyticsStore}, \ts{hordeGameStore}, \ldots):
spread the result into a store object literal; domain methods mutate
their fields and call \ts{this.notify()}. \ts{subscribe} returns the
matching unsubscribe function, so components can hook in directly or via
\ts{useSyncExternalStore}.
\end{apientry}

% =====================================================================
\chapter{Peak Finder Module (BattleShips)}
% =====================================================================
\label{ch:api-battleships}

All paths relative to \srcfile{modules/BattleShips/}.

\section{Type contracts (\srcfile{types/})}

\begin{apientry}{ProblemInstance (worked example)}{types/map.ts}
The central abstraction: game modes only ever hold a \ts{ProblemInstance},
never a raw \ts{MapConfig}.
\begin{lstlisting}
interface ProblemInstance {
  evaluate: (x: number, y: number) => number; // 0 = global min, 1 = far away
  bounds:   { xMin: number; xMax: number; yMin: number; yMax: number };
  isWin:    (x: number, y: number) => boolean;
  metadata?: { name?: string;
               description?: string;
               globalMinimum?: { x: number; y: number; value: number };
               winRadius?: number }; // extent of the win zone, for overlays
}
\end{lstlisting}
\apifield{Notes}Implemented by \ts{createMapProblem}
(\srcfile{engine/functionSurface.ts}) and \ts{createFunctionProblem}
(\srcfile{engine/functionProblem.ts}). New problem sources only need to
produce this shape.
\end{apientry}

\begin{apientry}{Coordinate, Minimum, MapConfig, ProbeResult}{types/map.ts}
\begin{lstlisting}
interface Coordinate { x: number; y: number }  // normalized 0-1
interface Minimum {
  id: string;
  position: Coordinate;
  isGlobal: boolean;  // exactly one minimum per map is the global one
  floor?: number;     // explicit depth of a local minimum; see Notes
}
interface MapConfig {
  id: string;               // 6-char uppercase share id
  minima: Minimum[];
  bounds: { xMin: number; xMax: number; yMin: number; yMax: number };
  winRadius: number;        // normalized radius around the global minimum
  createdAt: number;        // epoch ms
}
interface ProbeResult { position: Coordinate; value: number; isWin: boolean }
\end{lstlisting}
\apifield{Notes}\ts{Minimum.floor} is the surface value added on top of
the distance term in \ts{createMapProblem} --- lower = deeper, more
deceptive. When omitted, a position-seeded pseudo-random floor in
$[\ts{LOCAL\_MIN\_FLOOR\_MIN}, \ts{LOCAL\_MIN\_FLOOR\_MAX}] = [0.03, 0.1]$
is used, so the same map code always yields the same floors. The floor is
ignored (always 0) for the global minimum. \ts{bounds} is always the unit
square today.
\end{apientry}

\begin{apientry}{EAConfig (worked example)}{types/ea.ts}
Complete configuration of the Vs-EA opponent. \ts{DEFAULT\_EA\_CONFIG}
provides the defaults (identical to the \ts{'normal'} preset);
\ts{EA\_PRESETS} bundles named presets for the UI.
\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  \textbf{Field} & \textbf{Default} & \textbf{Meaning} \\
  \midrule
  \ts{populationSize}    & 30   & Individuals per generation \\
  \ts{maxGenerations}    & 200  & Hard stop \\
  \ts{crossoverRate}     & 0.7  & Probability of recombination \\
  \ts{mutationRate}      & 0.3  & Probability of mutating a child \\
  \ts{mutationStrength}  & 0.25 & Max.\ perturbation (normalized coords) \\
  \ts{mutationDecay}     & 0.90 & Per-generation strength multiplier \\
  \ts{winPopulationFraction} & 0.35 & Fraction inside win radius $\Rightarrow$ solved \\
  \ts{selectionStrategy} & \ts{'elitist'} & \ts{| 'tournament' | 'roulette'} \\
  \ts{crossoverStrategy} & \ts{'uniform'} & \ts{| 'arithmetic' | 'singlePoint'} \\
  \ts{mutationStrategy}  & \ts{'gaussian'}   & \ts{| 'uniform' | 'cauchy'} \\
  \bottomrule
\end{tabularx}
\end{center}
\end{apientry}

\begin{apientry}{Individual, Generation, SelectionFn, CrossoverFn,
MutationFn, RNG}{types/ea.ts}
\begin{lstlisting}
interface Individual {
  position:   Coordinate;
  fitness:    number;  // problem.evaluate(position) - minimized, 0 = summit
  isSolution: boolean; // inside the win radius
}
interface Generation {
  index:       number;
  individuals: Individual[]; // sorted best-first
  best:        Individual;
  meanFitness: number;
}
type RNG         = () => number; // Math.random-compatible
type SelectionFn = (population: Individual[], rng: RNG) => Individual;
type CrossoverFn = (a: Individual, b: Individual, rng: RNG) => Coordinate;
type MutationFn  = (coord: Coordinate, generationIndex: number,
                    config: EAConfig, rng: RNG) => Coordinate;
\end{lstlisting}
\apifield{Notes}Operator implementations are looked up from the strategy
registries in \srcfile{engine/ea/operators.ts}. Mutation receives the
generation index so it can apply the decayed strength
\ts{mutationStrength}\,$\cdot$\,\ts{mutationDecay}$^{\,gen}$.
\end{apientry}

\section{Engine (\srcfile{engine/})}

\begin{apientry}{createMapProblem}{engine/functionSurface.ts}
\ts{createMapProblem(config: MapConfig): ProblemInstance} --- builds the
smooth surface from the placed minima. The value at $(x, y)$ is
\[
  v(x, y) \;=\; \min\Bigl(1,\;\; \min_{m\,\in\,\mathrm{minima}}
  \bigl(d\bigl((x,y),\, m\bigr) + \mathit{floor}_m\bigr) \,/\, \ts{DISTANCE\_SCALE}\Bigr),
\]
where $\mathit{floor}_m$ is 0 for the global minimum and the explicit or
position-seeded floor for local minima --- so only the global minimum can
reach the perfect reading 0. \ts{isWin} tests
\ts{isWithinRadius(point, global, config.winRadius)}.
\apifield{Notes}Tunables: \ts{DISTANCE\_SCALE} (0.7; lower = values rise
faster around minima), \ts{LOCAL\_MIN\_FLOOR\_MIN/MAX} (0.03/0.1). Helper
exports: \ts{defaultLocalFloor(position)} (the seeded floor an untouched
node gets --- shown in the create UI), \ts{effectiveFloor(minimum)}, and
\ts{floorToCenterValue(floor)} (the value a probe reads exactly at a
minimum's center).
\end{apientry}

\begin{apientry}{functionProblem}{engine/functionProblem.ts}
Analytic benchmark surfaces (Sphere, Rastrigin, Eggholder, \ldots) behind
the same \ts{ProblemInstance} contract as hand-built maps.
\begin{lstlisting}
interface FunctionSpec {          // fully defines a playable surface
  fn: string;                     // benchmark id, or 'procedural'
  cx: number; cy: number;         // where the optimum lands in [0,1]^2
  theta: number;                  // rotation (radians)
  sx: number; sy: number;         // anisotropic scale (zoom)
  rx: 1 | -1; ry: 1 | -1;         // reflections
  seed?: number;                  // procedural surfaces only
}
createFunctionProblem(spec, sharpenOverride?): ProblemInstance
randomFunctionSpec(rng?, { id?, category? }): FunctionSpec
randomSurfaceSpec(): FunctionSpec      // sentinel: re-randomise every load
resolveSpec(spec, rng?): FunctionSpec  // sentinel -> concrete procedural
proceduralFunction(seed): BenchmarkFn  // seeded fBm "Mystery Surface"
encodeFunctionCode(spec): string / decodeFunctionCode(code): FunctionSpec
\end{lstlisting}
\ts{BENCHMARK\_FUNCTIONS} holds $\sim$20 classic test functions (several
BBOB-derived) grouped into \ts{FUNCTION\_CATEGORIES}
(\ts{simple | normal | complex | quirky}). \ts{createFunctionProblem}
grid-samples the affinely transformed surface once (64$\times$64) to
normalise values to $[0,1]$ (85th-percentile ceiling, so corner extremes
don't crush the colour range), applies a per-function sharpening exponent,
and pins the win target to the lowest sampled point. Winning is
\emph{value}-based: \ts{isWin} fires once \ts{evaluate} $\leq$
\ts{WIN\_VALUE\_EPS} (0.04), not at an exact coordinate;
\ts{metadata.winRadius} reports the measured extent of that zone for
replay overlays. Everything is deterministic in the spec (procedural
surfaces in \ts{seed}), so the main thread and the EA worker rebuild
identical surfaces.
\end{apientry}

\begin{apientry}{buildContours}{engine/contours.ts}
\begin{lstlisting}
function buildContours(
  evaluate: (x: number, y: number) => number, // surface, values in [0,1]
  levels: number[],                           // iso-values to trace
  resolution: number = 80,                    // grid density
): ContourLevel[]
\end{lstlisting}
Marching-squares isoline extraction. Samples the surface once on a
\ts{resolution}\,$\times$\,\ts{resolution} grid over $[0,1]^2$, then
emits, per iso-level, the line segments where the surface crosses that
level.
\apifield{Returns}One \ts{ContourLevel} (\ts{level} + \ts{segments}) per
requested level; each \ts{ContourSegment} is \ts{x1, y1, x2, y2} in the
same normalized $[0,1]^2$ space.
\end{apientry}

\begin{apientry}{geometry helpers}{engine/geometry.ts}
\begin{lstlisting}
euclideanDistance(a: Coordinate, b: Coordinate): number
isWithinRadius(point: Coordinate, center: Coordinate, radius: number): boolean
closestMinimumDistance(point: Coordinate, minima: Coordinate[]): number
\end{lstlisting}
Plain 2-D geometry on normalized coordinates. \ts{closestMinimumDistance}
returns \ts{Infinity} for an empty minima list (used by create mode's
spacing check).
\end{apientry}

\begin{apientry}{sampleGradient / sampleGradientRgb}{engine/colorScale.ts}
\ts{sampleGradient(t: number): [r, g, b]} --- shared colour ramp used by
heatmap, contours and charts: a full-spectrum gradient from deep violet
($t = 0$, the summit) through blue, cyan, green, yellow and orange to deep
red ($t = 1$, the valley), defined by an easily editable stop list.
\ts{sampleGradientRgb(t)} returns the same colour as a CSS
\ts{rgb()} string.
\end{apientry}

\begin{apientry}{valueToHeight}{engine/height.ts}
\ts{valueToHeight(value: number): number} --- presentation-only inversion
$1 - \ts{value}$ (clamped to $[0,1]$). Internally 0 = summit and 1 =
valley; the player is shown ``height'' with 1 = summit. Use only for
numbers displayed to the player --- never feed the result into colour
sampling or EA logic.
\end{apientry}

\begin{apientry}{encodeMap / decodeMap / generateRandomMap}{engine/mapCodec.ts}
\begin{lstlisting}
encodeMap(config: MapConfig): string  // URL-safe base64 share code
decodeMap(code: string): MapConfig    // throws Error('Invalid map code')
generateRandomMap(numMinima: number = 5): MapConfig
\end{lstlisting}
Map share-codes; format specification in Chapter~\ref{ch:api-formats}.
\ts{decodeMap} rejects unknown format versions and any unparsable input
with a single \ts{Invalid map code} error; missing \ts{wr}/\ts{t} fields
fall back to 0.04 / \ts{Date.now()}. \ts{generateRandomMap} places
\ts{numMinima} minima uniformly in $[0.05, 0.95]^2$ and promotes a random
one to global.
\end{apientry}

\begin{apientry}{problemCode}{engine/problemCode.ts}
\begin{lstlisting}
type ProblemSource =              // plain data - survives postMessage
  | { kind: 'map'; config: MapConfig }
  | { kind: 'fn';  spec: FunctionSpec };
decodeProblem(code: string): DecodedProblem  // throws on invalid codes
decodeProblemOrNull(code?: string): DecodedProblem | null
buildProblemFromSource(source: ProblemSource): ProblemInstance
\end{lstlisting}
Uniform entry point for \emph{any} PeakFinder code: function codes carry a
\ts{k:'fn'} marker and are tried first; anything else is treated as a map
code. \ts{DecodedProblem} bundles the rebuilt \ts{problem}, a display
\ts{label} (``Map \#ABC123'' / ``f Rastrigin''), the serializable
\ts{source} for the EA worker, and \ts{winTarget}/\ts{winRadius} for
overlays. A ``random every time'' sentinel spec is resolved once here on
the main thread, so the worker rebuilds the same concrete surface for the
whole session.
\end{apientry}

\begin{apientry}{codeClipboard}{engine/codeClipboard.ts}
\ts{copyCode(text): Promise<void>} / \ts{pasteCode(): Promise<string>} ---
clipboard helpers with an in-app fallback: every copy also lands in
\ts{localStorage} (\ts{bs.lastCopiedCode}), and paste falls back to that
value when the system clipboard is unreadable (Firefox web pages,
insecure origins).
\end{apientry}

\section{EA engine (\srcfile{engine/ea/})}

\begin{apientry}{createEAStepper (worked example)}{engine/ea/evolutionaryAlgorithm.ts}
\begin{lstlisting}
function createEAStepper(
  problem: ProblemInstance,
  config:  EAConfig,
  seed?:   number,          // seeds the LCG rng - same seed, same run
): EAStepper

interface EAStepper {
  step: (n: number) => StepResult;  // n = 0 reports without advancing
  currentGenIndex: () => number;
}
type StepResult =
  | { type: 'generation'; generation: Generation; replay?: ReplayFrame[] }
  | { type: 'solved'; generation: Generation; totalGenerations: number;
      replay?: ReplayFrame[] }
  | { type: 'exhausted'; totalGenerations: number; best: Individual };
\end{lstlisting}
Creates a resumable EA. The generation loop: sort best-first $\to$ copy
elite (top 5\%, $\geq 1$) $\to$ select parents $\to$ crossover (gated by
\ts{crossoverRate}) $\to$ mutate (gated by \ts{mutationRate}) $\to$ clamp
to bounds $\to$ evaluate $\to$ win check. The EA counts as solved once
$\max\bigl(2,\, \lceil \ts{populationSize} \cdot
\ts{winPopulationFraction} \rceil\bigr)$ individuals sit inside the win
radius simultaneously.
\apifield{Notes}The \ts{replay} on each result holds the frames of the
breeding that \emph{produced} the reported generation, so the replay's
final frame matches that generation's dots. \ts{runEA(problem, config,
callbacks, seed?)} is the non-interactive wrapper that steps to
completion, reporting through \ts{EACallbacks} (\ts{onGeneration} /
\ts{onSolved} / \ts{onExhausted}). The module-level
\ts{WIN\_POPULATION\_FRACTION} (0.10) is only a fallback for configs
missing \ts{winPopulationFraction}.
\end{apientry}

\begin{apientry}{SELECTION\_STRATEGIES, CROSSOVER\_STRATEGIES,
MUTATION\_STRATEGIES}{engine/ea/operators.ts}
Strategy registries mapping each \ts{*Strategy} string to its
implementation:
\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  & \textbf{Strategy} & \textbf{Behaviour} \\
  \midrule
  selection & \ts{tournament}  & Draw $k = 3$ random candidates, return the fittest. \\
            & \ts{roulette}    & Fitness-proportionate with inverted weights
                                 (\ts{maxFit} $-$ \ts{fitness}), so lower
                                 fitness wins more often. \\
            & \ts{elitist}     & Uniform pick from the top 20\% only. \\
  crossover & \ts{uniform}     & Each gene ($x$, $y$) drawn independently from either parent. \\
            & \ts{arithmetic}  & Blend $\alpha A + (1-\alpha) B$ with random $\alpha$. \\
            & \ts{singlePoint} & One parent contributes $x$, the other $y$. \\
  mutation  & \ts{gaussian}    & Normally-distributed offset (Box--Muller). \\
            & \ts{uniform}     & Flat offset within $\pm$strength. \\
            & \ts{cauchy}      & Heavy-tailed offset --- occasional long
                                 jumps help escape local minima. \\
  \bottomrule
\end{tabularx}
\end{center}
All mutation offsets are scaled by the decayed strength
\ts{mutationStrength}\,$\cdot$\,\ts{mutationDecay}$^{\,gen}$.
\apifield{Notes}\ts{CROSSOVER\_STRATEGIES\_RECORDING} is the single source
of truth: those variants additionally return the random choices made
(\ts{CrossoverResult} --- blend factor $\alpha$ / per-gene source), which
the replay log needs and which cannot be reconstructed from the child
position once mutation has been applied. The plain
\ts{CROSSOVER\_STRATEGIES} are thin coordinate-only wrappers that draw
identical RNG values.
\end{apientry}

\begin{apientry}{individual helpers}{engine/ea/individual.ts}
\begin{lstlisting}
createRandom(problem, rng): Individual  // uniform in problem.bounds
evaluate(position, problem): Individual // wraps fitness + isSolution
clamp(coord, problem): Coordinate       // clip to problem.bounds
\end{lstlisting}
\end{apientry}

\begin{apientry}{eaReplayLog}{engine/ea/eaReplayLog.ts}
\ts{buildReplayFrames(before, nextGen, eliteCount, config,
solutionThreshold, breeding?)} builds the frame list for one generation;
\ts{snapshot(ind, id)} freezes an individual with a stable per-replay id.
\ts{ReplayFrame} is a discriminated union on \ts{phase}; every frame
carries \ts{headline} and \ts{description} (ready-to-display prose) plus
the population as \ts{IndividualSnapshot[]}:
\begin{center}
\begin{tabularx}{\textwidth}{l X}
  \toprule
  \textbf{Phase} & \textbf{Extra payload} \\
  \midrule
  \ts{sorted}    & population ranked best-first \\
  \ts{elite}     & \ts{eliteIds}; \ts{nextGen} (the elites copied so far) \\
  \ts{selection} & \ts{strategy}; tournament: \ts{candidateIds},
                   \ts{winnerId}; roulette: normalized \ts{weights};
                   elitist: \ts{eliteTopCount} \\
  \ts{breeding}  & \ts{parentAId}/\ts{parentBId}, \ts{child},
                   \ts{crossoverResult} (pre-mutation position),
                   \ts{crossoverStrategy}, \ts{alpha}/\ts{geneSource},
                   \ts{didCrossover} \\
  \ts{mutating}  & \ts{childId}, \ts{didMutate},
                   \ts{before}/\ts{after}/\ts{delta} \\
  \ts{newGen}    & the completed next generation \\
  \ts{winCheck}  & \ts{solutionIds}, \ts{count}, \ts{threshold}, \ts{solved} \\
  \bottomrule
\end{tabularx}
\end{center}
\apifield{Notes}The breeding/mutating frames show the \emph{real} first
non-elite child: a \ts{BreedingRecord} (parents, gate outcomes,
pre-mutation position) is captured live inside the stepper's breeding
loop, because those random choices cannot be reconstructed afterwards.
The union also declares an \ts{initial} phase, but \ts{buildReplayFrames}
never emits it --- the stepper hands over an already-sorted population.
\end{apientry}

\section{Worker protocol (\srcfile{engine/ea/ea.worker.ts})}

\begin{apientry}{WorkerInMessage / WorkerOutMessage (worked example)}{types/ea.ts}
The EA runs in a dedicated Web Worker
(\srcfile{engine/ea/ea.worker.ts}); one stepper persists in worker scope
between messages. The main thread sends \ts{WorkerInMessage}s and receives
\ts{WorkerOutMessage}s:

\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  \textbf{Direction} & \textbf{Type} & \textbf{Payload and meaning} \\
  \midrule
  in  & \ts{START} & \ts{config: EAConfig}, \ts{source: ProblemSource} ---
        rebuilds the problem and creates the stepper. Does \emph{not}
        step, and posts nothing back (generation 0 arrives with the first
        \ts{STEP}, avoiding a duplicate). \\
  in  & \ts{STEP}  & \ts{count: number} --- advance up to \ts{count}
        generations. \\
  in  & \ts{STOP}  & Discard the stepper. \\
  out & \ts{GENERATION} & \ts{generation: Generation},
        \ts{replay?: ReplayFrame[]} --- the replay of the breeding that
        produced this generation rides along on the same message. \\
  out & \ts{SOLVED} & \ts{generation}, \ts{totalGenerations},
        \ts{replay?}. \\
  out & \ts{EXHAUSTED} & \ts{totalGenerations}, \ts{best: Individual} ---
        \ts{maxGenerations} reached without solving. \\
  out & \ts{ERROR} & \ts{message: string} --- e.g.\ the problem could not
        be rebuilt from \ts{source}. \\
  \bottomrule
\end{tabularx}
\end{center}
\end{apientry}

\section{Hooks (\srcfile{hooks/})}

\begin{apientry}{useEARunner (worked example)}{hooks/useEARunner.ts}
\begin{lstlisting}
const ea = useEARunner();
ea.init(source, eaConfig); // create worker, send START - does NOT run
ea.step(n);                // advance n generations (default 1)
ea.stop(); ea.reset();
// state: ea.status, ea.generations, ea.currentGeneration, ea.best,
//        ea.totalGenerations, ea.latestReplay, ea.errorMessage
\end{lstlisting}
React-side owner of the EA worker. \ts{source} is a serializable
\ts{ProblemSource}; \ts{config} defaults to \ts{DEFAULT\_EA\_CONFIG}.
\ts{EAStatus = 'idle' | 'running' | 'solved' | 'exhausted' | 'error'}.
\ts{latestReplay} holds the \ts{ReplayFrame[]} of the most recent step
batch, consumed by the replay overlay.
\apifield{Notes}Once solved, \ts{totalGenerations} is locked to the
solving generation --- further probes keep the worker stepping, but no
longer advance the ``solved in $X$ generations'' tally. \ts{start} is
currently an alias of \ts{init}. Worker errors terminate the worker and
surface in \ts{errorMessage}.
\end{apientry}

\begin{apientry}{replay playback}{src/hooks/useReplayPlayer.ts}
There is no PeakFinder-specific replay hook: \ts{EAReplayOverlay} feeds
its \ts{ReplayFrame[]} to the shared, frame-agnostic
\ts{useReplayPlayer<Frame>(frames, autoplayIntervalMs?)}, which provides
\ts{currentFrame}, \ts{frameIndex}, \ts{totalFrames},
\ts{isFirst}/\ts{isLast}, \ts{next}/\ts{prev}/\ts{goTo} and
\ts{play}/\ts{pause} autoplay.
\end{apientry}

\begin{apientry}{useGameMap}{hooks/useGameMap.ts}
Create-mode map state.
\begin{lstlisting}
const map = useGameMap(maxMinima?);
// state:   minima, mapId, maxMinima, minSpacing, isFull, hasGlobal,
//          globalMinimum
// actions: addMinimum(pos), removeMinimum(id), setGlobalMinimum(id),
//          setFloor(id, floor | undefined), clearAll(),
//          getMapConfig(), getCode()
\end{lstlisting}
\ts{addMinimum} silently rejects placements beyond \ts{MAX\_MINIMA} (12)
or closer than \ts{MIN\_SPACING} (0.12) to an existing minimum.
\ts{setFloor(id, undefined)} reverts a node to its random default depth.
\ts{clearAll} also mints a fresh \ts{mapId}, so each created map saves as
its own entry. \ts{getCode()} = \ts{encodeMap(getMapConfig())}; the win
radius is fixed at 0.04.
\end{apientry}

\begin{apientry}{usePlaySession}{hooks/usePlaySession.ts}
\begin{lstlisting}
const { probes, status, bestProbe, probe, reset } =
  usePlaySession(problem: ProblemInstance | null);
\end{lstlisting}
Play-mode session state. \ts{probe(position)} evaluates the problem and
appends a \ts{ProbeResult}; \ts{status} is
\ts{'idle' | 'playing' | 'won'} and stays \ts{'won'} once won (probing
remains possible for post-win exploration). \ts{bestProbe} is the probe
with the lowest value so far.
\end{apientry}

\begin{apientry}{useSavedMaps}{hooks/useSavedMaps.ts}
\begin{lstlisting}
const { savedMaps, saveMap, removeMap, renameMap } = useSavedMaps();
interface SavedMap { id; name?; code; createdAt; minimaCount }
\end{lstlisting}
The player's saved-map library --- a thin domain wrapper around the
shared \ts{useSavedLibrary} hook, persisted in \ts{localStorage} under
\ts{bs.savedMaps}. \ts{saveMap(config: MapConfig)} stores the encoded
share code; saving an id that already exists is a no-op.
\end{apientry}

\section{Key components (\srcfile{components/game/})}

\begin{apientry}{GameMap (worked example)}{components/game/shared/GameMap.tsx}
The central square map canvas (grid + heatmap + minima dots + overlays).
\begin{lstlisting}
minima?: Minimum[]; showMinima?: boolean; highlightGlobal?: boolean;
evaluateFn?: (x, y) => number;  // enables the HeatmapLayer
heatmapConfig?: Partial<HeatmapConfig>;
revealPoints?: Coordinate[];    // heatmap only shown near these
exclusionRadius?: number;       // dashed rings around minima (create mode)
onMapClick?: (coord: Coordinate) => void;  // normalized coords
onMinimumClick?: (id: string) => void; selectedId?: string | null;
overlayLabel?: string; className?: string; children?: ReactNode;
\end{lstlisting}
Clicks are converted to normalized $[0,1]^2$ coordinates; clicks on a
minimum dot go to \ts{onMinimumClick} instead of \ts{onMapClick}. Probe
markers, win rings etc.\ are passed as absolutely-positioned
\ts{children}.
\end{apientry}

\begin{apientry}{ContourLayer / HeatmapLayer}{components/game/shared/}
Both render a surface visualization from
\ts{\{ evaluate, config?, revealPoints? \}}. Reveal semantics:
\ts{undefined} reveals the whole map; an array (even an empty one)
reveals only circles of \ts{revealRadius} around its points. \ts{GameMap}
currently renders the heatmap; \ts{ContourLayer} (SVG marching-squares
isolines, levels packed densely near the summit) remains available for
contour-style views. Defaults: \ts{DEFAULT\_CONTOUR\_CONFIG} =
\ts{\{ lineCount: 24, spacingExponent: 2.5, resolution: 100,
revealRadius: 0.1 \}}; \ts{DEFAULT\_HEATMAP\_CONFIG} =
\ts{\{ resolution: 1080, opacity: 0.82, revealRadius: 0.05,
valueExponent: 0.55 \}}.
\end{apientry}

\begin{apientry}{EAReplayOverlay / ReplayMap / IndividualList}
{components/game/vs-ea/}
Full-screen replay modal: \ts{EAReplayOverlay(\{ frames, onClose \})}
drives the shared \ts{useReplayPlayer} through the frame list.
\ts{ReplayMap} renders the population dots plus per-phase decorations:
\ts{highlightIds}/\ts{dimIds}/\ts{markerIds}/\ts{solutionIds} sets,
\ts{customOpacities} (roulette weights), \ts{extraDots} (the bred child),
\ts{arrow} (mutation offset), \ts{parentLine} (arithmetic crossover) and
\ts{geneLines} (uniform/single-point gene sources). Dots glide with
\ts{DOT\_MOVE\_DURATION} = 1000\,ms. \ts{IndividualList} lists the
population with role badges (\ts{ELITE}, \ts{PARENT A/B}, \ts{CHILD},
\ts{WIN}, \ts{WINNER}, \ts{CANDIDATE}, \ts{ELIGIBLE}) via
\ts{roles: Map<id, IndividualRole>} and optional per-row
\ts{annotations}.
\end{apientry}

\begin{apientry}{FitnessChart, ModeSelector, CodeModal}{components/game/shared/}
\ts{FitnessChart(\{ series, yMax?, compact?, yLabel?, hoveredIndex?,
onHover?, markers? \})} --- SVG line chart of probe/EA values over time
(y-axis shown as the player-facing ``height'').
\ts{ModeSelector(\{ onSelect \})} --- the Create / Play / Vs~EA landing
picker. \ts{CodeModal(\{ code, mapId, onClose, onPlayNow? \})} ---
share-code dialog with clipboard copy via \ts{copyCode}.
\end{apientry}

\begin{apientry}{SavedMapsSidebar / SavedFunctionsSidebar}{components/game/shared/}
Two self-contained left-edge drawers (no props).
\ts{SavedMapsSidebar} lists the player's saved maps (\ts{useSavedMaps})
with rename/delete; clicking an entry copies its share code.
\ts{SavedFunctionsSidebar} lists the analytic benchmark functions grouped
by category; copying a function mints a \emph{fresh} random transform
each time, and ``Random every time'' copies the sentinel spec that
re-randomises the whole surface on every load.
\end{apientry}

\begin{apientry}{VsEAMode and overlays}{components/game/vs-ea/}
\ts{VsEAMode} wires the race: it owns the \ts{EAConfig} (initialised to
the \ts{'normal'} preset), steps the EA per player probe, and mounts
\ts{EASettingsPanel} (difficulty presets from \ts{EA\_PRESETS} plus
advanced sliders), \ts{EAWinOverlay} / \ts{SecondSolveOverlay} (EA won /
player beat its solution on the retry) and
\ts{GenerationReplayOverlay} (dot movement across whole generations, as
opposed to the phase-by-phase \ts{EAReplayOverlay}).
\end{apientry}

% =====================================================================
\chapter{Maze Explorer Module (mazeGame)}
% =====================================================================
\label{ch:api-maze}

All paths relative to \srcfile{modules/mazeGame/}. The module mirrors the
BattleShips structure with a \emph{discrete} genome: an individual is a
fixed-length move sequence instead of a point.

\section{Type contracts (\srcfile{types/})}

\begin{apientry}{Maze core types (worked example)}{types/maze.ts}
\begin{lstlisting}
interface Cell { x: number; y: number }  // grid coordinates
type Move = 0 | 1 | 2 | 3;          // indexes into MOVE_ARROWS / MOVE_DELTAS
type Path = Move[];                  // the genome: fixed-length move sequence
const MOVE_ARROWS = ['↑','→','↓','←'];
type WallRule  = 'waste' | 'break' | 'repair';
type FitnessFnId = 'manhattan' | 'geodesic' | 'length' | 'novelty';

interface Grid { cols: number; rows: number; walls: Uint8Array }
interface SerializedMaze extends Grid { start: Cell; goal: Cell }
interface WalkResult {          // the phenotype of a Path
  visited: Set<number>;         // visited cell indices (y*cols + x)
  trail: Cell[];                // cell per step; length = path.length + 1
  finalCell: number;            // where the agent settled
  reachedGoalAt: number;        // step it first stood on the goal, or -1
  crashedAt: number;            // wall-crash step ('break' rule only), or -1
}
interface MazeProblem {
  grid: Grid; cols; rows; start: Cell; goal: Cell;
  geodesicField: Int32Array;    // BFS distance to goal; -1 = unreachable
  pathLength: number;           // auto-sized genome length L
  fitnessFnId: FitnessFnId; wallRule: WallRule;
  evaluate: (walk: WalkResult) => number;  // lower = better, 0 = optimal
  isWin: (walk: WalkResult) => boolean;
  metadata: { seed: number; shortestPath: number };
}
\end{lstlisting}
\apifield{Notes}\ts{Grid.walls[i]} holds an \emph{open-passage} bitmask
per cell $i = y \cdot \ts{cols} + x$ (N=1, E=2, S=4, W=8; a set bit means
the passage is open --- \ts{MOVE\_WALL\_BIT}); \ts{cellIndex(x, y, cols)}
maps 2-D coordinates to flat indices. \ts{SerializedMaze} decouples a
hand-built creator maze from procedural generation (start/goal are
arbitrary). \ts{MazeProblem} parallels BattleShips'
\ts{ProblemInstance}, but \ts{evaluate}/\ts{isWin} take a simulated
\ts{WalkResult} instead of an $(x, y)$ coordinate. \ts{WallRule} is the
constraint-handling strategy for blocked moves: \ts{waste} loses the
move, \ts{break} crashes and freezes the agent (death penalty),
\ts{repair} rewrites the gene to a random open direction and takes it
(Lamarckian).
\end{apientry}

\begin{apientry}{EAConfig / DEFAULT\_MAZE\_EA\_CONFIG}{types/ea.ts}
Discrete-EA configuration. Shape mirrors the BattleShips \ts{EAConfig},
plus three maze-only fields (\ts{fitnessFnId}, \ts{wallRule},
\ts{pathLengthFactor}); an \ts{Individual} carries a \ts{path} and its
cached \ts{walk} instead of a position.
\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  \textbf{Field} & \textbf{Default} & \textbf{Meaning} \\
  \midrule
  \ts{populationSize}    & 100  & Individuals per generation \\
  \ts{maxGenerations}    & 300  & Hard stop \\
  \ts{crossoverRate}     & 0.85 & Probability of recombination \\
  \ts{mutationRate}      & 0.95 & Probability of mutating a child \\
  \ts{mutationStrength}  & 0.06 & Per-gene re-pick \emph{probability} (not a
                                  perturbation magnitude) \\
  \ts{mutationDecay}     & 1.0  & Per-generation strength multiplier (no
                                  decay --- keeps exploration alive) \\
  \ts{winPopulationFraction} & 0.08 & Fraction that must reach the goal
                                  $\Rightarrow$ solved \\
  \ts{selectionStrategy} & \ts{'tournament'} & \ts{| 'roulette' | 'elitist'} \\
  \ts{crossoverStrategy} & \ts{'singlePoint'} & \ts{| 'uniform'} \\
  \ts{mutationStrategy}  & \ts{'point'} & \ts{| 'segment'} \\
  \ts{fitnessFnId}       & \ts{'geodesic'} & Which fitness function the
                                  problem evaluates against \\
  \ts{wallRule}          & \ts{'waste'} & Blocked-move handling
                                  (\ts{| 'break' | 'repair'}) \\
  \ts{pathLengthFactor}  & 3.5  & Genome length as a multiple of the
                                  shortest path ($\geq 1$; baked in at
                                  build time --- changing it restarts the
                                  run) \\
  \bottomrule
\end{tabularx}
\end{center}
\end{apientry}

\section{Engine (\srcfile{engine/})}

\begin{apientry}{mazeGen}{engine/mazeGen.ts}
\begin{lstlisting}
generateMaze(cols, rows, seed, { braid?, openness? }): Grid
pickRandomStartGoal(grid, rng, minFrac = 0.6): { start; goal }
gridFromEdgeWalls(cols, rows, isHWall, isVWall): Grid
\end{lstlisting}
Seeded procedural generation --- identical seed $\Rightarrow$ identical
maze (the basis of the fitness-comparison experiment). \ts{generateMaze}
runs an iterative recursive backtracker (every cell reachable), then two
seeded post-passes that only ever \emph{remove} walls, so connectivity is
preserved: \ts{braid} opens a fraction of dead-ends into loops,
\ts{openness} removes a fraction of the remaining interior walls
(punching side doors into the backtracker's long corridors, creating
parallel routes the population can split across).
\apifield{Notes}\ts{pickRandomStartGoal} places the goal anywhere and
draws the start from cells whose corridor distance to the goal is at
least \ts{minFrac} of the maze's BFS diameter --- never trivially close.
\ts{gridFromEdgeWalls} converts the maze creator's edge-wall model into a
\ts{Grid}. \ts{mazeToAscii} is a debug-only ASCII renderer.
\end{apientry}

\begin{apientry}{geodesic}{engine/geodesic.ts}
\begin{lstlisting}
computeGeodesic(grid: Grid, goal: Cell, start: Cell): GeodesicResult
interface GeodesicResult {
  field: Int32Array;         // corridor distance to goal; -1 = unreachable
  diameter: number;          // largest finite distance in the field
  shortestFromStart: number; // start -> goal distance
}
\end{lstlisting}
BFS distance field flooded outward from the goal through open corridors,
computed once at problem build time. Backs the \ts{geodesic} and
\ts{length} fitness functions, genome auto-sizing, and
\ts{pickRandomStartGoal}.
\end{apientry}

\begin{apientry}{mazeProblem}{engine/mazeProblem.ts}
\ts{createMazeProblem(opts: BuildMazeOptions): MazeProblem} --- generates
the maze (or accepts an explicit \ts{grid} for creator mazes), floods the
geodesic field, auto-sizes the genome length
$L = \min\bigl(\ts{MAX\_PATH\_LENGTH},\, \lceil \mathrm{shortest} \cdot
\ts{pathLengthFactor} \rceil\bigr)$, and wires up the chosen fitness
function. Changing only \ts{fitnessFnId} for the same seed yields the
\emph{same} maze with a different \ts{evaluate} --- the comparison demo.
All fitness functions are ``lower = better'', normalized to roughly
$[0,1]$, and score the agent's \emph{final} cell (the goal absorbs), so a
lucky fly-through earns no credit:
\begin{center}
\begin{tabularx}{\textwidth}{l X}
  \toprule
  \textbf{Id} & \textbf{Scoring} \\
  \midrule
  \ts{manhattan} & Straight-line distance of the final cell to the goal
    --- ignores walls, so it is \emph{deceptive}: the population piles up
    in the dead-end nearest the goal as the crow flies, and stalls. \\
  \ts{geodesic}  & Corridor (BFS) distance of the final cell to the goal,
    normalized by the maze diameter --- smooth and solvable. \\
  \ts{length}    & Geodesic plus a reward for reaching the goal early:
    non-solvers score in $[0.5, 1]$, solvers in $[0, 0.5)$ scaled by
    \ts{reachedGoalAt} --- reaching dominates, shorter wins among
    solvers. \\
  \ts{novelty}   & Placeholder (constant 1): novelty depends on the whole
    population + archive, so the stepper applies a \ts{NoveltyScorer}
    that overwrites each individual's fitness. \\
  \bottomrule
\end{tabularx}
\end{center}
\apifield{Notes}\ts{isWin} is \ts{reachedGoalAt >= 0} regardless of
fitness function. Defaults: \ts{DEFAULT\_BRAID} = 0.5,
\ts{DEFAULT\_OPENNESS} = 0.25 (too high and Manhattan stops being
deceptive), \ts{MAX\_PATH\_LENGTH} = 600 (exported so the creator can
warn when a maze's shortest path exceeds what a genome can hold).
\end{apientry}

\begin{apientry}{createNoveltyScorer (worked example)}{engine/ea/novelty.ts}
\begin{lstlisting}
function createNoveltyScorer(
  cols: number, rows: number, k: number = DEFAULT_K,
): NoveltyScorer
\end{lstlisting}
Novelty-search infrastructure --- the canonical maze-deception demo
(Lehman \& Stanley): objective-chasing gets trapped, pure exploration
floods the maze and stumbles onto the exit. Instead of rewarding
closeness to the goal, each individual is rewarded for ending up
somewhere \emph{few} others (now or in the past) have ended up. Used when
\ts{FitnessFnId = 'novelty'}; because it depends on the whole population
plus an archive, the stepper calls \ts{scorer.score(pop)} each generation
instead of a per-individual \ts{evaluate}.
\apifield{Notes}Behaviour descriptor = the final cell the agent settles
on. \ts{score} sets each individual's fitness to
$1 - \min(1,\, \mathrm{novelty} / \mathrm{maxDist})$, where novelty is
the mean Euclidean distance to the $k$ nearest behaviours ($k = 15$ by
default, own behaviour excluded) among the current population plus the
archive. Archive admission: every \emph{newly explored} distinct final
cell is added (deduplicated by cell index); \ts{archiveSize()} exposes
the count for visualization.
\end{apientry}

\begin{apientry}{individual helpers / walkPath}{engine/ea/individual.ts}
\begin{lstlisting}
walkPath(problem, path, rng?): WalkResult   // simulate the phenotype
evaluate(path, problem, rng?): Individual   // walk + score + isSolution
createRandom(problem, rng): Individual      // random genome of length L
repair(path, problem, rng): Path            // enforce genome length L
randomMove(rng): Move
\end{lstlisting}
\ts{walkPath} executes the genome move by move. The trail is always
\ts{path.length + 1} cells (uniform for animation): a blocked move keeps
the agent in place under \ts{waste}, freezes it for the rest of the walk
under \ts{break} (recorded in \ts{crashedAt}), or rewrites the gene to a
random open direction under \ts{repair} (Lamarckian --- \ts{evaluate}
therefore copies the genome first, so shared parent arrays are never
mutated). The goal \emph{absorbs}: once reached, the agent stays there,
which is what makes the final cell a meaningful fitness target.
\ts{repair} is the length-invariant safety net (truncate / pad with
random moves).
\end{apientry}

\begin{apientry}{SELECTION\_STRATEGIES, CROSSOVER\_STRATEGIES\_RECORDING,
MUTATION\_STRATEGIES}{engine/ea/operators.ts}
Selection is ported verbatim from BattleShips (tournament $k = 3$ /
roulette with inverted weights / elitist top 20\% --- it only reads
\ts{.fitness}). Crossover and mutation work on the move sequence:
\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  & \textbf{Strategy} & \textbf{Behaviour} \\
  \midrule
  crossover & \ts{singlePoint} & Follow parent A up to a random cut
              (1..$L{-}1$), then switch to parent B; records \ts{cut}. \\
            & \ts{uniform}     & Each gene independently drawn from either
              parent; records the per-gene \ts{geneMask}. \\
  mutation  & \ts{point}       & Each gene independently re-picked with
              probability \ts{mutationStrength}\,$\cdot$\,%
              \ts{mutationDecay}$^{\,gen}$. \\
            & \ts{segment}     & Re-randomize one contiguous block whose
              max length scales with the decayed strength (a ``kink'' the
              path forks at). \\
  \bottomrule
\end{tabularx}
\end{center}
\apifield{Notes}Both operator families are recording variants
(\ts{CrossoverResult} / \ts{MutationResult} with \ts{mutatedIndices}) ---
the replay log needs the splice cut / gene mask / mutated positions,
which cannot be reverse-engineered from the child genome.
\end{apientry}

\begin{apientry}{createMazeEAStepper}{engine/ea/evolutionaryAlgorithm.ts}
\ts{createMazeEAStepper(problem, config, seed?): EAStepper} --- same
resumable-stepper contract and \ts{StepResult} union as BattleShips
(Chapter~\ref{ch:api-battleships}), with two maze-specific additions:
\begin{itemize}[nosep]
  \item \ts{updateConfig(nextConfig, nextProblem?)} --- swap in new tuning
        \emph{without} discarding the population. When \ts{nextProblem}
        is given (fitness function or wall rule changed), every genome is
        kept but re-walked and re-scored under the new problem.
  \item Novelty runs use a solution threshold of 1 (a single explorer
        finding the exit counts as solved) --- novelty spreads the
        population out, so it never piles a fraction onto the goal. All
        other fitness functions use
        $\max(2, \lceil \ts{populationSize} \cdot
        \ts{winPopulationFraction} \rceil)$.
\end{itemize}
Elitism is top 5\% ($\geq 1$), as in BattleShips.
\end{apientry}

\begin{apientry}{eaReplayLog}{engine/ea/eaReplayLog.ts}
Same phase set and frame structure as the BattleShips replay log
(\ts{sorted} $\to$ \ts{elite} $\to$ \ts{selection} $\to$ \ts{breeding}
$\to$ \ts{mutating} $\to$ \ts{newGen} $\to$ \ts{winCheck}), re-based on
paths: \ts{IndividualSnapshot} carries \ts{path}, \ts{trail} and
\ts{finalCell} instead of a position (the map draws trails as spaghetti,
the list renders genomes as arrow strings). Payload differences:
\ts{breeding} records \ts{cut}/\ts{geneMask} instead of
$\alpha$/\ts{geneSource}; \ts{mutating} records
\ts{beforePath}/\ts{afterPath}/\ts{mutatedIndices} instead of a
coordinate delta.
\end{apientry}

\begin{apientry}{mazeCodec}{engine/mazeCodec.ts}
\begin{lstlisting}
mazeId(maze: SerializedMaze): string   // content-derived 6-char id
encodeMaze(maze: SerializedMaze): string
decodeMaze(code: string): SerializedMaze // throws Error('Invalid maze code')
\end{lstlisting}
Maze share-codes; format in Chapter~\ref{ch:api-formats}. \ts{mazeId}
hashes the maze content (djb2 $\to$ base36), so identical mazes get
identical ids --- what the saved-maze library de-duplicates on.
\ts{decodeMaze} validates version, dimensions (2..64 per axis), wall
string length and marker bounds, and throws a single
\ts{Invalid maze code} error otherwise.
\end{apientry}

\section{Worker protocol (\srcfile{engine/ea/maze.worker.ts})}

\begin{apientry}{WorkerInMessage / WorkerOutMessage}{types/ea.ts}
Same shape as the BattleShips protocol; the out-messages
(\ts{GENERATION} / \ts{SOLVED} / \ts{EXHAUSTED} / \ts{ERROR}) are
identical apart from the maze-typed payloads. Differences on the in-side:

\begin{center}
\begin{tabularx}{\textwidth}{l l X}
  \toprule
  \textbf{Direction} & \textbf{Type} & \textbf{Payload and meaning} \\
  \midrule
  in & \ts{START} & \ts{config: EAConfig}, \ts{seed}, \ts{cols},
       \ts{rows}, \ts{maze?: SerializedMaze} --- a hand-built creator
       maze takes precedence; otherwise the maze is generated procedurally
       from \ts{seed}. The EA RNG is seeded with the same value, so on the
       same seed the \emph{only} thing that varies between fitness
       functions is \ts{evaluate} --- the comparison demo. \\
  in & \ts{STEP} & \ts{count: number}. \\
  in & \ts{UPDATE\_CONFIG} & \ts{config: EAConfig} --- retune the running
       EA without restarting; takes effect on the next generation. Only a
       changed \ts{fitnessFnId} or \ts{wallRule} rebuilds the problem
       (and re-scores the population); everything else is pure tuning. \\
  in & \ts{STOP} & Discard the stepper. \\
  \bottomrule
\end{tabularx}
\end{center}
\apifield{Notes}\ts{pathLengthFactor} is fixed for the worker's lifetime:
the main thread sends a fresh \ts{START} when it changes, so mixed-length
populations can never occur.
\end{apientry}

\section{Hooks and components}

\begin{apientry}{useMazeEARunner}{hooks/useMazeEARunner.ts}
\begin{lstlisting}
const ea = useMazeEARunner();
ea.init({ seed, cols, rows, maze?, pathLengthFactor? }, config?);
ea.step(n); ea.updateConfig(config); ea.stop(); ea.reset();
// state: status, generations, currentGeneration, best,
//        totalGenerations, latestReplay, errorMessage
\end{lstlisting}
Mirrors BattleShips' \ts{useEARunner} (same \ts{EAState}, same
solved-lock on \ts{totalGenerations}, \ts{start} again an alias of
\ts{init}); the differences are the \ts{START} payload
(\ts{MazeRunParams} instead of a \ts{ProblemSource}), the default config
(\ts{DEFAULT\_MAZE\_EA\_CONFIG}) and the extra \ts{updateConfig} for
live retuning. Replay playback again uses the shared
\ts{useReplayPlayer} --- there is no maze-specific replay hook.
\end{apientry}

\begin{apientry}{useSavedMazes}{hooks/useSavedMazes.ts}
\begin{lstlisting}
const { savedMazes, saveMaze, removeMaze, renameMaze } = useSavedMazes();
interface SavedMaze { id; name?; code; createdAt; cols; rows }
\end{lstlisting}
Saved-maze library over the shared \ts{useSavedLibrary}, persisted in
\ts{localStorage} under \ts{maze.savedMazes}. Entries are keyed by the
content-derived \ts{mazeId}, so saving an identical maze is a no-op;
\ts{saveMaze} returns the id either way.
\end{apientry}

\begin{apientry}{MazeCanvas}{components/shared/MazeCanvas.tsx}
The shared maze renderer (walls, start/goal markers, overlays). Key
props: \ts{cols}/\ts{rows}/\ts{walls}/\ts{start}/\ts{goal} plus
\ts{trails} (population / player paths, drawn in order),
\ts{solidCells} (creator's painted blocks), \ts{markers} (highlight dots,
e.g.\ splice / mutated cells), \ts{agents} (animated dots drawn above
walls), \ts{previewWall} (creator hover ghost), \ts{showGrid}, and a
fog-of-war cell set used by Solve mode to reveal the maze as it is
explored.
\end{apientry}

\begin{apientry}{MazeSetupScreen / MazeExperimentMode /
MazeEASettingsPanel}{components/experiment/}
\ts{MazeSetupScreen} --- experiment entry: pick a fresh random maze (size
/ openness of your choosing) or a saved creation, then start.
\ts{MazeExperimentMode} --- the run itself: owns the \ts{EAConfig},
drives \ts{useMazeEARunner}, and pins settings changes to the generation
where they took effect as markers on the fitness chart.
\ts{MazeEASettingsPanel(\{ config, onConfigChange, shortestPath?,
runStarted?, onClose \})} --- the \ts{EAConfig} editors (also exported
chrome-free as \ts{MazeEASettingsControls}); the genome-length slider is
locked once a run has started, since length is baked into every genome.
\end{apientry}

\begin{apientry}{MazeEAReplayOverlay / MazePathMap / MazeIndividualList}
{components/experiment/replay/}
Full-screen dissection of one generation --- a re-skin of BattleShips'
replay overlay with an identical phase set.
\ts{MazeEAReplayOverlay(\{ frames, problem, onClose \})} steps the shared
\ts{useReplayPlayer}. \ts{MazePathMap} draws every individual's walked
trail over the maze (highlight/dim/marker/solution id-sets,
\ts{customOpacities} for roulette, \ts{selectedId} to lift one path above
the rest, splice-point cell markers). \ts{MazeIndividualList} renders
genomes as arrow strings with role badges, per-gene highlights
(crossover source / mutated genes), \ts{maxGenes} truncation and
click-to-select (\ts{onSelect}).
\end{apientry}

\begin{apientry}{MazeCreateMode / MazePlayMode}{components/create/, components/play/}
\ts{MazeCreateMode} --- the maze editor: draw/erase walls on cell edges
(elliptical edge hitboxes so free-form strokes don't grab stray
orthogonal walls), place start/goal, size 4--32 per axis, export via
\ts{encodeMaze}. \ts{MazePlayMode} --- the player writes a move string
with a D-pad and watches it walk (one step per 140\,ms tick), edits it,
and re-submits --- experiencing genome, phenotype and mutation by hand;
a ghost replays the previous attempt for comparison.
\end{apientry}

% =====================================================================
\chapter{Shooter Module (shooterGame)}
% =====================================================================
\label{ch:api-shooter}

All paths relative to \srcfile{modules/shooterGame/}.

\section{Types and constants (\srcfile{shooter.types.ts})}

\begin{apientry}{DNA (worked example)}{shooter.types.ts}
An agent genome is \ts{number[]} of length~7; all genes in $[0,1]$.
Indices via \ts{DNA\_INDEX}:
\begin{center}
\begin{tabularx}{\textwidth}{c l X}
  \toprule
  \# & \textbf{Gene} & \textbf{Effect} \\
  \midrule
  0 & \ts{AGGRESSION}      & Chase strength \\
  1 & \ts{DODGE\_WEIGHT}   & Bullet-avoidance strength \\
  2 & \ts{SHOOT\_ACCURACY} & Aim precision (1 = perfect) \\
  3 & \ts{PREFERRED\_RANGE}& Combat distance ($\times$300\,px + 100\,px) \\
  4 & \ts{MOVEMENT\_SPEED} & Speed bonus (40 base + gene $\times$ 80\,px/s) \\
  5 & \ts{PREDICT\_LEAD}   & Aim lead factor \\
  6 & \ts{FIRE\_RATE}      & Fire rate (higher = faster) \\
  \bottomrule
\end{tabularx}
\end{center}
Related constants: \ts{ARENA} (800$\times$800 logical px),
\ts{GAME\_CONFIG} (round duration, bullet speed, cooldowns),
\ts{STARTER\_DNA}. \ts{GamePhase = 'idle' | 'playing' | 'roundEnd' |
'evolving'}.
\end{apientry}

\section{Core loop (\srcfile{game/core/})}

\begin{apientry}{gameLoop.update}{game/core/gameLoop.ts}
Pure function \ts{update(state, dt, input) → GameState}; no React
dependencies. Also accumulates \ts{PlayerGhostFrame[]} into
\ts{GameState.ghostFrames} for ghost-based evolution.
\todo{Full signature and state transitions.}
\end{apientry}

\begin{apientry}{renderer}{game/core/renderer.ts}
Canvas draw calls. \todo{Entry.}
\end{apientry}

\begin{apientry}{vec}{game/core/vec.ts}
2-D vector math: \ts{add}, \ts{sub}, \ts{scale}, \ts{normalize},
\ts{angle}, \ts{clamp}, \ts{perpendicular}, \ldots \todo{Signatures.}
\end{apientry}

\section{Genetic algorithm (\srcfile{game/ga/})}

\begin{apientry}{evolve (worked example)}{game/ga/evolution.ts}
Produces the next generation after a round: tournament selection $\to$
single-point crossover $\to$ per-gene mutation, driven by each agent's
round fitness. Companions:
\begin{itemize}[nosep]
  \item \ts{getNextAgent(population): number[]} --- DNA of the next
        opponent to spawn;
  \item \ts{presimulate(generations): Population} --- round-robin
        pre-training (every agent vs.\ every other) before the first round;
  \item \ts{presimulateAgainstGhost(generations, ghost)} --- evolves the
        population against a recording of the player's previous round.
\end{itemize}
\todo{Exact signatures and config knobs.}
\end{apientry}

\begin{apientry}{fitness}{game/ga/fitness.ts}
\ts{calculateFitness(stats: RoundStats): number},
\ts{calculateRaidbossFitness(stats, roundDuration)},
\ts{normalizeFitness(fitnesses)}. \todo{Scoring formulas.}
\end{apientry}

\begin{apientry}{population}{game/ga/population.ts}
\ts{initPopulation}, \ts{updatePopulationStats}. \todo{Entry.}
\end{apientry}

\begin{apientry}{evolution.worker.ts}{game/ga/evolution.worker.ts}
\todo{Worker message protocol for off-thread evolution/presimulation.}
\end{apientry}

\section{Stores}

The shooter uses custom observable stores (not Zustand/Redux): components
subscribe via \ts{store.subscribe(cb)} and read \ts{store.state} directly.

\begin{apientry}{gameStore}{game/gameStore.ts}
Holds \ts{GameState} outside React; consumed by \ts{ShooterCanvas} and
\ts{DNADisplay}. \todo{State shape and mutation API.}
\end{apientry}

\begin{apientry}{raidbossStore / analyticsStore}{game/}
\todo{Entries; document the Firebase interaction of the raidboss store.}
\end{apientry}

\section{Horde mode (\srcfile{horde/})}

\begin{apientry}{horde subsystem}{horde/}
\ts{hordeTypes.ts}, \ts{hordeDna.ts}, \ts{hordeMaps.ts},
\ts{hordePathfinding.ts}, \ts{hordeCollision.ts}, \ts{hordeGameStore.ts},
\ts{hordeRunStore.ts}. \todo{One entry per file: map format, pathfinding
algorithm, wave/DNA evolution, store shapes.}
\end{apientry}

\section{Mods (\srcfile{mods/})}

\begin{apientry}{mods subsystem}{mods/}
\ts{modTypes.ts}, \ts{runModsStore.ts}, \ts{shotEngine.ts}.
\todo{Mod type registry and how mods alter the shot engine.}
\end{apientry}

\section{Hooks and components}

\begin{apientry}{useGameLoop / useInput / useTouchControls}{hooks/}
\ts{useGameLoop}: requestAnimationFrame loop calling \ts{update()} and
driving \ts{gameStore}. \ts{useInput}: keyboard + mouse $\to$
\ts{InputState}. \ts{useTouchControls}: virtual joystick / aim zone.
\todo{Signatures.}
\end{apientry}

\begin{apientry}{ShooterCanvas, HordeCanvas, DNADisplay,
MobileJoystickZone, MobileAimZone}{components/}
\todo{One short entry each.}
\end{apientry}

% =====================================================================
\chapter{Data Formats}
% =====================================================================
\label{ch:api-formats}

Serialization formats are the closest thing the app has to a public API:
they cross device boundaries via share codes and local storage.

\section{Peak Finder map code (v1)}

Produced by \ts{encodeMap} (\srcfile{BattleShips/engine/mapCodec.ts}):
a URL-safe Base64 encoding of

\begin{lstlisting}
{ "v": 1,                      // format version
  "id": "ABC123",              // map id
  "m": [[x, y, isGlobal, floor?], ...],  // minima (normalized coords)
  "wr": 0.04,                  // win radius (fraction of map width)
  "t": 1699999999999 }         // timestamp
\end{lstlisting}

The optional 4th tuple entry (\ts{floor}) is only present for local minima
with an explicit depth --- omitting it keeps older codes valid and
round-trips untouched (random-floor) nodes unchanged. Coordinates are
rounded to 4 decimals. \ts{decodeMap} throws a single
\ts{Error('Invalid map code')} for anything unparsable or with an unknown
\ts{v}; a missing \ts{wr} defaults to 0.04 and a missing \ts{t} to
\ts{Date.now()}.

\section{Maze code}

Produced by \ts{encodeMaze} in \srcfile{mazeGame/engine/mazeCodec.ts}: a
URL-safe Base64 encoding of

\begin{lstlisting}
{ "v": 1,             // format version
  "id": "ABC123",     // content-derived id (djb2 hash, base36)
  "c": cols, "r": rows,
  "s": [x, y],        // start cell
  "g": [x, y],        // goal cell
  "w": "af3...",      // one hex digit per cell: open-passage nibble
                      // (N=1 E=2 S=4 W=8), row-major
  "t": 1699999999999 }// timestamp
\end{lstlisting}

The full wall grid is encoded (not a generator seed), so hand-built
creator mazes round-trip exactly; 32$\times$32 --- the creator's maximum
--- is 1024 hex digits, comfortably copy-pasteable. \ts{decodeMaze}
validates version, dimensions (2..64), wall-string length and marker
bounds, throwing \ts{Error('Invalid maze code')} otherwise.

\section{Function problem code}

Produced by \ts{encodeFunctionCode} in
\srcfile{BattleShips/engine/functionProblem.ts}: a URL-safe Base64
encoding of

\begin{lstlisting}
{ "k": "fn",              // marker - distinguishes from map codes
  "fn": "rastrigin",      // benchmark id, "procedural", or "random"
  "c": [cx, cy],          // optimum position in the [0,1] viewport
  "th": theta,            // rotation (radians)
  "s": [sx, sy],          // anisotropic scale
  "r": [rx, ry],          // reflections (1 | -1)
  "sd": seed }            // only for procedural surfaces
\end{lstlisting}

\ts{decodeProblem} (\srcfile{BattleShips/engine/problemCode.ts}) tries the
function decoder first (keyed on the \ts{k:"fn"} marker), then falls back
to map decoding. A code with \ts{fn:"random"} is a sentinel that resolves
to a fresh procedural surface on every load.

\section{Local storage keys}

\begin{center}
\begin{tabularx}{\textwidth}{l X}
  \toprule
  \textbf{Key} & \textbf{Contents} \\
  \midrule
  \ts{bs.hints.enabled} & Hint system on/off (\ts{localStorage}) \\
  \ts{bs.hints.seen}    & Seen hint ids (\ts{sessionStorage}, per session) \\
  \ts{bs.savedMaps}     & Saved Peak Finder maps (\ts{useSavedMaps}) \\
  \ts{bs.lastCopiedCode}& In-app clipboard fallback for share codes \\
  \ts{maze.savedMazes}  & Saved mazes (\ts{useSavedMazes}) \\
  \bottomrule
\end{tabularx}
\end{center}

\appendix

% =====================================================================
\chapter{Extension Recipes}
% =====================================================================

Short how-tos for the most likely extension points.
\todo{Flesh out each recipe with concrete steps.}

\begin{itemize}
  \item \textbf{Add a new analytic problem to Peak Finder} --- implement a
        \ts{ProblemInstance} via \ts{createFunctionProblem} and register it.
  \item \textbf{Add a new EA operator} --- add an entry to the relevant
        \ts{*\_STRATEGIES} registry and extend the strategy union type.
  \item \textbf{Add a new maze fitness function} --- extend
        \ts{FitnessFnId}, \ts{FITNESS\_FN\_IDS} and \ts{FITNESS\_FN\_LABELS},
        implement the scorer in \ts{mazeProblem.ts}.
  \item \textbf{Add a help topic} --- one entry in \ts{HELP\_TOPICS} +
        \ts{<HelpButton topic="..."/>}.
  \item \textbf{Add a hint} --- one entry in \ts{hintContent.ts}, fire via
        \ts{showHint(id)} or wrap with \ts{<HintPopover>}.
  \item \textbf{Add a new mini-game module} --- follow the four-layer module
        structure (Chapter~2) and register a route in \srcfile{App.tsx}.
\end{itemize}

% =====================================================================
\chapter{Tunable Constants Reference}
% =====================================================================

Central table of gameplay/visual constants and their locations.
\todo{Import the full tunables table (currently maintained in
\srcfile{claude.md}) and keep the two in sync — or declare one canonical.}

\end{document}
